---
title: "Tutorial to Use the HAT Package in R"
output: rmarkdown::html_vignette
bibliography: citation.bib
vignette: >
  %\VignetteIndexEntry{Tutorial to Use the HAT Package in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Simeng Shao
#### February 25, 2021

The `hat` package implements a hierarchical aggregation testing algorithm, as described in the paper "Controlling the False Split Rate in Tree-Based Aggregation". The package implements a hierarchical testing function that determines how to aggregate leaves with while controlling the *false split rate* (FSR), an error measure defined in the above paper.  There are two primary use cases considered in the paper:

1. aggregating observations with the same means and

2. aggregating features with the same coefficients in linear regression.

We provide examples of these two use cases and then demonstrate a more general example of the hierarchical aggregation procedure.

* [Application to observation aggregation](#observation)
* [Application to feature aggregation](#feature)
* [General example](#general)

<a id="observation"></a>


## Application to observation aggregation

### Brief look at the data

This section is an example of using the package to achieve aggregation of observations based on their means. We calculate volatility of daily stocks price data that are derived from the US Stock Database ©2021 Center for Research in Security Prices (CRSP), The University of Chicago Booth School of Business @CRSP.

The tree structure is constructed based on the North American Industry Classification System ([NAICS](https://www.census.gov/naics/)), an industry classification system that employs a six digit code: the first two digits designate the largest sector; the third, fourth, fifth and sixth digits designate the subsector, industry group, industry, and national industry, respectively. 

One can load the volatility data and pre-trained tree structure into our environment:

```{r, eval=T, include=T}
library(hat)
nrow(stocks_volatility)
colnames(stocks_volatility)
```

The `stocks_volatility` data set contains `r nrow(stocks_volatility)` stocks' corresponding company names and the volatility during the 5-year period.

The distribution of volatility is right-skewed. Therefore, in our analysis, we will take the logarithm to reduce skew.

The tree structure is saved in `stocks_tree`, which is formed according to the NAICS hierarchy. We can look at the tree structure by using the function `plot_hc_list()`. However, we omit this for now because we will see the tree again very soon.

### Aggregating observations

The idea of aggregation is based on the fact that some companies might share the same mean volatility of stock if they are "similar enough" in the tree. We assume a model

\[
y_i = \theta_{k(i)} + \epsilon_i, \ \ \ k(i)\in\{1,...,K\}, i\in \{1, ..., p\},
\]

where, in the stocks example, $y_i$ is the volatility of the $i$-th stock. We will use the tree as a guide as it describes the similarity among companies and therefore helps narrow down reasonable aggregations to consider.

Our algorithm achieves aggregation in two main steps:

1. Generate a p-value by an ANOVA test for each interior node.

2. Use the hierarchical aggregation testing procedure to sequentially go down the tree in a fashion that will control the FSR of the overall aggregation of leaves that is produced.

The function `aggregate_observations` performs both of these steps:

```{r, eval=T, include=T}
result = aggregate_observations(y = log(stocks_volatility$volatility), 
                       sigma_constant = NULL, 
                       hc_list = stocks_tree, 
                       alpha_level = 0.4)
```

* `sigma_constant`: standard deviation of noise $\epsilon$'s, if known.  If `sigma_constant` is non-`NULL`, then a chi-squared null distribution is used; if `sigma_constant` is `NULL`, then an F-test is performed.

* `hc_list`: a list of length-$|\mathcal{T}\setminus \mathcal{L}|$ that describes the tree structure. This format is a generalization of the `merge` matrix within an `hclust` object.  The element `hc_list[[i]]` contains the children of the $i$-th node on the tree. A negative value indicates that the child is a leaf node while positive values are interior nodes. For example, if `hc_list[[i]][j]` is positive, that means that node `hc_list[[i]]` is a parent of  node `hc_list[[j]]`.  The object `stocks_tree` in our package is an example of such a structure. The function can also work with `hclust` objects (for binary trees) or `dendrogram` objects (for more general trees).  Only one of the three is required.  The functions `dend_as_hclist` and `hclist_as_dend` allow one to switch between formats.

* `alpha_level`: target FSR level.

The output `result` is a list that has three components: `alpha` is the target FSR level, `groups` gives the group assignments of the observations, and `rejections` indicates if each interior node is rejected (this will come in handy for plotting the aggregation result on a tree).

In this example, the achieved aggregation result contains `r length(unique(result$groups))` groups. We plot the achieved aggregation on the tree using the function `plot_aggregation` (a bit busy as since are >2k stocks...).

```{r, fig.height=4, fig.width=7, eval=T, include=T}
plot_aggregation(rejections = result$rejections, hc_list = stocks_tree)
```


<a id="feature"></a>

## Application to feature aggregation

```{r, eval=T, include=F}
set.seed(1234)
n = 100
p = 50
k = 10
hc = hclust(dist((1:p) + runif(p)/p), method = "complete")
groups = cutree(hc, k)
coeffs = runif(k, 0, 18)[groups]
X = matrix(rnorm(p * n), nrow = n, ncol = p)
X = X * matrix(rbinom(n*p, size = 1, prob = 0.6),nrow = n, ncol = p)
y = X %*% coeffs + rnorm(n)
```

For this application, we will synthetize data for feature aggregating. We randomly generate a tree with `r p` leaves by hierarchical clustering, and cut the tree into `r k` disjoint groups. Then we simulate `r k` coefficient values corresponding to the `r k` groups.
To mimick the scenario of rare features (@Yan2018RareFS), we generate a design matrix $X$ by a Gaussian-Bernoulli distribution.
Finally, we build a linear model with gaussian noise as the response vector $y$.

To perform aggregation, we can use the function `aggregate_features`. The input requires 

- `y`: response vector.

- `X`: design matrix.

- `hc`: an object that describes the tree structure. Since here we have a binary tree, we use a `hc = hclust` object. Alternatively, one can also use `dend = dendrogram` or `hc_list = hc_list` if they have a non-binary tree that is saved in the format of 'dendrogram' or 'hc_list'.

- `alpha_level`: target FSR level.

```{r, eval=T, include=T}
agg_result = aggregate_features(y = y, X = X,
                          hc = hc, alpha_level = 0.2)
```

For calculating FSP we can use 

```{r, eval=T, include=T}
calculate_fsp(theta_est = agg_result$groups, theta_true = groups)
```

<a id="general"></a>

## General example

The functions `aggregate_observations` and `aggregate_features` shown above both make use of the hierarchical aggregation testing framework introduced in the paper "Controlling the False Split Rate in Tree-Based Aggregation".  Internally, both functions call a more general function called `hierarchical_testing`.  If a user has a different way to construct nodewise p-values for aggregation, then the user should use `hierarchical_testing` as shown here.

This function takes a length-`num_interior_node` vector of p-values.

Here is a small example:

```{r}
set.seed(123)
n <- 20
hc <- hclust(dist(runif(n)))
de <- as.dendrogram(hc)
hc_list <- dend_as_hclist(de)

# compute_my_pvalue <- function(u) {
#   # Example of a customized pvalue function:
#   # u an element of hc_list
#   
#   # compute something more interesting than just runif here.
#   
#   # hc_list[[u]]
# }
# 
# result_small <- hierarchical_test(hc_list = hc_list, 
#                   p_vals = lapply(hc_list, compute_my_pvalue),
#                   alpha_level = 0.4, 
#                   independent = FALSE)
# 
# plot_aggregation(rejections = result_small$rejections, hc_list = hc_list)
```




### References




